### 2018-10-17 paper.tex
Spelling and grammar revised up to end of Section II

\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{relsize}
\usepackage{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\usepackage{array}
%\usepackage{caption}
%\usepackage[caption=false,font=footnotesize]{subfig}
%\usepackage{fixltx2e}
\usepackage{float}
\usepackage{url}
%\usepackage{becs}
\usepackage{listings}
\lstset{%
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  stepnumber=1,
  xleftmargin=2em
}
\usepackage{color}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\let\latexvec=\vec
\let\vec=\latexvec
\interdisplaylinepenalty=2500
\newtheorem{definition}{Definition}
\newtheorem{exampleb}{Example}
\newcommand{\subsubsubsection}[1]{\medskip\par\textit{#1:}}
\newcommand{\card}[1]{\vert{#1}\vert}

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}

\title{A New Metric for the Analysis of Swarms using Potential Fields}
\author{\uppercase{Neil Eliot}\authorrefmark{1},
\uppercase{David Kendall\authorrefmark{1}, and Michael Brockway\authorrefmark{1}}}
\address[1]{Northumbria University, Department of Computing and Information Sciences, 
Newcastle upon Tyne, NE1 8ST}

\markboth
{Eliot \headeretal: A New Metric for the Analysis of Swarms using Potential Fields}
{Eliot \headeretal: A New Metric for the Analysis of Swarms using Potential Fields}

\corresp{Corresponding author: Neil Eliot (e-mail: neil.eliot@northumbria.ac.uk).}

\begin{abstract}
  % An important criterion in the assessment of swarm coordination algorithms is
  % the extent to which they cause unnecessary movement, i.e. movement which
  % occurs as a side effect of the algorithm rather than from the fulfilment of
  % the intended function of the swarm. 
  There are many metrics defined for the analysis of swarm coordination
  algorithms. These metrics are usually based upon the distances between
  agents, the distance between agents and a fixed point, or the resultant vectors
  that \textit{potential field} effects produce. This paper examines a
  distance-based metric that measures a swarm's overall structure using
  inter-agent distances. It introduces a new metric that identifies a swarm's
  state based upon the resultant magnitude of the vectors produced by the
  agent interactions that create the agent distribution within the swarm's
  structure. The algorithms used to implement the swarming feature are based
  upon cohesion and repulsion vectors between an agent and its neighbours. In
  comparing and contrasting the two metrics, we find that the
  \textit{cohesion/repulsion} metric offers a number of advantages over the
  distance metric. In particular, the \textit{cohesion/repulsion} metric allows
  the identification of the essential characteristic of a swarm as `expanding',
  `stable', or `contracting'. These states cannot be identified using a
  distance-based metric. Practical swarming applications where the new metric
  can be applied advantageously include area-filling and reconnaissance.
\end{abstract}

\begin{keywords}
Swarming, Swarm Dynamics, Swarm Metric.
\end{keywords}

\titlepgskip=-15pt
\maketitle

\section{Introduction\label{sec:intro}}

%\section{Introduction\label{methods:SwarmStability}}

\textit{Swarming} in the animal kingdom of ants, bees, fish and birds, for
instance, has long been studied by scientists. From these studies, mathematical
models and algorithms have evolved, such as those produced by
Reynolds~\cite{REY:87}. The models and algorithms have in turn captured the
interest of computing scientists, who are interested in applying them to large
groups of autonomous mobile \textit{agents} (`robots'). The cooperative
coordination of these agents can take many forms, such as following a set
path~\cite{HCS:09}, existing in a static space~\cite{EP:10, GP:02, GP:04} or
foraging as a colony~\cite{HER:11, GK:07}. One of the attributes of swarms that
has captured the interest of scientists is that the models and algorithms used
to coordinate them are generally sets of simple rules. These simple rules cause
the agents to appear to work cooperatively. Often, the rules can also generate
movement that does not contribute to the goal of the swarm. This movement is a
direct result of the agents attempting to move towards an optimum position
within the structure. This internal movement can be identified by analysing the
changes in the inter-agent interactions. Measuring this unnecessary movement
allows the effectiveness of a swarming algorithm to be evaluated. If the
internal movement can be reduced, then the stability of the swarm is improved
and more of the swarm's energy is focused on the goal. Reducing the internal
movement also provides a more stable platform for the deployment of sensor
arrays. Both metrics, presented in this paper, allow this evaluation to be
carried out when all the sensor ranges are the same. However, the distance metric
is not able to identify the internal movement accurately when the sensor ranges
of each agent vary, producing irregular inter-agent distances, even though the
interactions at the vector level indicate that the agents are in fact stable
and optimum. This paper discusses the two metrics and how they differ in their
approach to identifying these changes. \textit{Distance-based} metrics use
variations in the inter-agent spaces, as discussed by many research teams
including Navarro et al.~\cite{NIM:09}, Gazi and Passino~\cite{GP:02,GP:04},
Barnes et al.~\cite{BAF:06} and Schneider and Wildermuth~\cite{SW:03}. The new
metric uses the resultant magnitudes of the agents' \textit{inter-agent
vectors} that are defined by interactions of an agent with its neighbours. This
allows the ``degree'' of a relationship to be evaluated globally and therefore
allows a swarm to be identified as ``expanding'', ``stable'' or
``contracting''.

Many coordination algorithms exist to control the movement of agents within a
swarm. The primary goal of these algorithms is to maintain a stable swarm
structure. A stable structure is one where the agents in the swarm ``act'' as a
single entity. The main focus of this paper is on metrics that can be
applied to any coordination algorithms that use \textit{potential fields} to
produce a swarming effect. Most swarming algorithms define a set of equations
that calculate the repulsion and cohesion effects. These equations are the
basis of the new metric.

To facilitate the analysis of the new metric, a basic swarm model is introduced
and then two metrics are applied to this model to present a comparison of the
information provided by a \textit{distance-based} and the new
\textit{cohesion/repulsion-based} metric. The analysis includes the
identification of a swarm's ``state''. The metric is then applied to an 
area-filling scenario to highlight a possible use of the new metric.

The remainder of this paper is structured as follows:
Section~\ref{Section:Related} discusses different metrics that are currently
used to analyse swarms; Section~\ref{Section:SwarmModelling} introduces a basic
swarm model to produce an environment that can be analysed and evaluated using
the two metrics; Section~\ref{Section:StabilityMagnitude} discusses the
mathematical model that produces the swarming effect and how the potential
fields are applied; Section~\ref{Section::Movement1} discusses how the model is
implemented over time to produce movement;
Section~\ref{Section:StabilityNullVector} discusses the effect of the model on
swarm stability and the effects models have on swarm movements; 
% Section~\ref{Section:Jitter} discusses the effect the mathematical model can
% have upon the structure of a swarm in terms of potentially wasted energy. 
Section~\ref{Section:DistanceDynamics} discusses a distance-based metric for
swarm analysis; Section~\ref{Section:MagnitudeDynamics} introduces the new
\textit{cohesion/repulsion-based} metric, showing how it is calculated and
applied; Section~\ref{Section:ApplicationFloodFilling} compares the two metrics
using an area-filling scenario;
Section~\ref{Section:MagnitudeDistanceComparison} discusses the results of
experiments, and, finally, Section~\ref{Section:FutureWork} discusses
additional work that needs to be carried out to exploit the additional
information that the new metric provides.

\section{Related Work}\label{Section:Related}

Many metrics exist for the analysis of swarms. These metrics can be categorised
as \textit{centroidal-based}, \textit{agent/inter-agent distance-based} and
\textit{magnitude-based}. The common denominator of \textit{centroidal} and
\textit{agent/inter-agent distance} metrics is that they \textit{measure} the
agent distribution within a swarm based upon distances.
\textit{Magnitude-based} metrics analyse a swarm based upon inter-agent
interaction.

\textit{Centroidal} analysis involves identifying the centroid of a swarm, also
referred to as the \textit{centre of mass}, as shown in Figure~\ref{com} as a
red cross and defined in Equation~\ref{eq:Centroid1}.

\begin{equation}\label{eq:Centroid1}
  \bar{r} = \frac{1}{\card{S}}\mathlarger{\sum_{b \in S}} r_b
\end{equation}

Here, $\bar{r}$ denotes the position vector of the swarm's \textit{centroid};
$r_b$ is the position of each agent, $b$, and $S$ is the set of agents that make
up the swarm.

Navarro et al. use the centroid to produce a "Mean Distance from the Boundary
to the Center of the Group" metric~\cite{NIM:09}. This metric takes the
distance of each agent from the centroid and calculates the mean and deviation
to produce their measure. Gazi and Passino use this technique when analysing
the stability of a swarm based on agent distribution~\cite{GP:11}. The
\textit{centroid} is also used as a reference point to determine the path of a
swarm. Gazi and Passino use the centroid when analysing the stability of a
swarm which is designed to maintain its position, by analysing the changes in
the centroid's position over time~\cite{GP:04}. Gazi and Passino also use a
centroid-based analysis in~\cite{GP:02}~to identify directional changes in
agent movement by creating a ``map'' of the potential fields of a swarm in its
entirety and overlaying that to the ``centre'' of the swarm.  Barnes et. al.
use the centroid to analyse the effects of using bivariate normal functions in
their work on using potential fields~\cite{BAF:06, BAFVM:06}.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4.5cm]{figures/com}
{Swarm Centre of Mass.\label{com}}

\textit{Inter-agent distance} analysis is a basic mechanism for analysing a
swarm in terms of separation of all of the agents with respect to each agent's
visible neighbours. This is different to using the \textit{centroidal} approach,
as the measure is based on agent interactions rather than a resultant structure
which affects the centroid. This technique is the most popular approach to
analysing a swarm's structure. Gazi and Passino use a distance-based analysis
in~\cite{GP:04, GP:11} to identify optimum distribution of agents based upon
different cohesion and repulsion applications. Wu et. al. discuss using
distance-based metrics to evaluate emergent behaviours in swarms based on a
Viscsek model~\cite{WCW:11}. These include calculating the average number of
neighbours for each agent, grouping agents based on numbers of neighbours (to
identify clusters), and calculating the average distance between agents~\cite{WCW:11}.
Schneider and Wildermuth~\cite{SW:03} use a slightly different approach to
measuring performance. They identify the difference in the paths of robots in a
simulated versus a physical implementation and calculate a distance variance
between them.

\textit{Magnitude-based} analysis uses the \textit{potential fields} to
calculate the ``degree'' of interaction between agents. Bennet and McInnes use
the resultant relationship between agents to identify equilibrium states when
creating structures such as rings and controlling the direction of agents to
create specific transformations~\cite{BM:09}. Wu et al. introduce the concept
of a \textit{rejection potential} in an attempt to generate a "kinetic energy"
score for a swarm~\cite{WCW:11}.

This paper builds upon the work relating to the \textit{magnitude-based} metric
analysis, as discussed by Bennet and McInnes, and Wu et al., by expanding the
analysis to the entire swarm and deriving a scalar value that can be used in the
analysis of swarm behaviour.

%% \section{Metric principles\label{section:MagnitudeDynamics}}
%% This section discusses the theory behind swarm movement and the effects that can be measured using both the metrics described above.
\section{Swarm modelling}\label{Section:SwarmModelling}
Currently, much swarm research uses field effects as the method of modelling
inter-agent interactions~\cite{BAF:06, BAFVM:06, BM:09, APZDAMC:09, GP:02,
GP:04, GP:04a, GP:05, GP:11, MYP:09}. The models usually use two field effects
to implement the swarming characteristic. These effects are \textit{cohesion},
to draw agents closer, and \textit{repulsion} to prevent agents colliding.
\textit{Potential fields} are ranges around an agent that determine the effect
other agents have upon its movement~(Fig.~\ref{methods:FieldEffects}). It is
usual for the cohesion field to have a radius $C_b$ which is larger than the
repulsion radius $R_b$. When an agent $b'$ moves into the \textit{cohesion
field} of an agent $b$, then $b'$ is said to be a neighbour of $b$, and $b$ is
subject to cohesion. When an agent $b'$ moves into the \textit{repulsion field}
of $b$, then $b$ is subject to repulsion. When the repulsion magnitude exceeds
the cohesion magnitude, the agent $b$ will have a tendency to move away from
$b'$, i.e. to be repelled. Obstacles can be introduced into the model. When an
agent $b$ moves too close to an obstacle, i.e. within the obstacle repulsion
range $O_b$, it has a \textit{repulsion vector} applied, giving it a tendency
to move away from the obstacle.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7.5cm]{figures/FieldEffects}
{Agent field effects.\label{methods:FieldEffects}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/FieldEffects}
%% \end{center}
%% \caption{Agent field effects\label{methods:FieldEffects}}
%% \end{figure}

\section{Inter-agent vector magnitude effect on internal movement}\label{Section:StabilityMagnitude}
Figure~\ref{methods:Stability5} shows the cohesion and repulsion vector
contributions of neighbour $b'$ to $v_c(b)$ and $v_r(b)$, as given in
Equations~\ref{eq:FlyToCentre1} and \ref{eq:Repulsion1}. Notice that the vectors
are along the line of separation $bb'$. \\

%% Equation (\ref{eq:Neighbours1}) identifies the set of agents that are neighbours of $b$.
%% 
%% \begin{equation}\label{eq:Neighbours1}
%% nbr(b) \buildrel \Delta \over = \{b' \in S | |bb'| <= C_b\}
%% \end{equation}‎

Equation~\ref{eq:FlyToCentre1} identifies the resultant cohesion effect of all
the neighbours of $b$. Here, $nbr(b)$ defines the set of all the agents in the
swarm that are a neighbour of $b$; a neighbour is any agent that falls within
the cohesion field.

\begin{equation}\label{eq:FlyToCentre1}
v_{c}(b) = \frac{1}{|nbr(b)|}\left({\mathlarger{\sum_{b' \in nbr(b)}}{bb'}}\right)
\end{equation}

Equation~\ref{eq:Repulsion1} identifies the resultant repulsion effect of all
the neighbours of $b$, where $rep(b)$ returns the set of all agents that are
within the repulsion field of $b$.

\begin{equation}
\label{eq:Repulsion1}
v_{r}(b) = 
\frac{1}{|rep(b)|}
\left(
\mathlarger{\mathlarger{\sum_{b' \in rep(b)}}}
{\left( 1-\frac{|bb'|}{R_b} \right)}
{bb'}
\right)
\end{equation}

Using the cohesion and repulsion vectors generated by the relationship of $b$
to its neighbours, a resultant vector can be calculated that summarises the
effects to which the agent is being subjected, and provides the direction in
which the agent will move, based on this relationship. This is the
\textit{inter-agent vector}.  Equation~\ref{eq:InterAgentMovement1} identifies
the \textit{inter-agent vector} for agent $b$ with respect to its neighbours.

\begin{equation}\label{eq:InterAgentMovement1}
v(b) = v_{c}(b) + v_{r}(b)
\end{equation}

If a \textit{weighted} model is used then~Equation~\ref{eq:InterAgentMovement1}
is replaced by~Equation~\ref{eq:InterAgentMovement2}, where $k_c$ and $k_r$ are
weightings to change the effect of each vector, as exemplified in
Table~\ref{tab:MetricPhysics1}.

\begin{equation}\label{eq:InterAgentMovement2}
v(b) = k_cv_{c}(b) + k_rv_{r}(b)
\end{equation}

%% From here throughout sections \ref{metric:MagnitudeDynamics2} and \ref{metric:StabilityNullVector} imagine agent $b$ has just a single neighbour $b'$ and consider the effect of $b'$ on $v(b)$, the \textit{inter-agent vector} of $b$.
%% The relationship between the agents is that: \textit{the vector for $b$ with respect to $b^{'}$, and the vector for $b^{'}$ with respect to $b$ are on the same line}~(\ref{methods:Stability5}).
%% If we consider an agent ($b$) as having only one neighbour ($b'$) then using the formulae defined in chapter~\ref{chapter:methods} the vectors $v_c(b)$~(\ref{eq:FlyToCentre1}) and $v_r(b)$~(\ref{eq:Repulsion1}) are the magnitudes of the vectors that are generated by the field effects between the two agents. $k_c$ and $k_r$ are the weighting factors for the cohesion and repulsion. The resulting vectors are aligned along the same line of separation between the agents.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=7cm]{figures/Stability5}
{Vectors on line of separation.\label{methods:Stability5}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=7cm]{figures/Stability5}
%% \end{center}
%% \caption{Vectors on line of separation} \label{methods:Stability5}
%% \end{figure}

\section{Internal movement and the null vector}\label{Section:StabilityNullVector}

When the two vectors (cohesion and repulsion) have magnitudes that are equal
and opposite they produce a null vector, indicating that two agents are
optimally spaced for a given set of conditions. Although the agents are at an
optimum position when resultants are zero it does not mean the swarm is
optimally distributed. If a swarm is in a confined space it is possible for an
optimum position to be created where the vector magnitude is affected by a
compression effect. This phenomenon is used in the identification of the
emergent behaviour of \textit{area filling}.

If we consider the equilibrium state (Fig.~\ref{methods:Stability4}) the
resultant $v(b)=(0,0)$. A null vector cannot be normalised to produce a
directional vector ($\hat{v} = \frac{v}{|v|}$ if $v\neq0$; $0$ if $v=0$) The
effect is that the agent will remain stationary. If all agent pairs are in this
condition the swarm will stop moving~(Fig.~\ref{methods:StabilityNullVector}).

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector}
{Equilibrium with null vectors.\label{methods:StabilityNullVector}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/StabilityNullVector}
%% \end{center}
%% \caption{Equilibrium with null vectors} \label{methods:StabilityNullVector}
%% \end{figure}

Due to the movement of agents being based only on neighbour interactions this
situation is very rare. The residual motion that persists in a swarm is the
background `noise' or `jitter' that an algorithm creates.

If a swarm is goal-based, the addition of a \textit{directional vector} will
prevent all agents simultaneously producing null vectors as shown
in~Fig.~\ref{concave:VoidPerimeter2} and
Equation~\ref{eq:InterAgentMovementDirectional1}. Here, $k_dv_{d}(b)$ is the
calculated directional vector to be applied to an agent.

\begin{equation}\label{eq:InterAgentMovementDirectional1}
v(b) = k_cv_{c}(b) + k_rv_{r}(b) + k_dv_{d}(b)
\end{equation}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector3}
{Equilibrium with directional vector (t).\label{concave:VoidPerimeter1}}

%% \begin{figure}[H]
%% \begin{center}
%% 	 \includegraphics[width=5.5cm]{figures/StabilityNullVector3}
%% \end{center}
%% \label{concave:VoidPerimeter1}
%% \caption{(t)}
%% \end{figure}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=4cm]{figures/StabilityNullVector2}
{Equilibrium with directional vector (t + 1).\label{concave:VoidPerimeter2}}

%% \begin{figure}[H]
%% \begin{center}
%% 	 \includegraphics[width=6cm]{figures/StabilityNullVector2}
%% \end{center}
%% \caption{(t + 1)}
%% \label{concave:VoidPerimeter2}
%% \end{figure}

% \section{Residual internal movement (Jitter)}\label{Section:Jitter}
% Due to the dynamic nature of a swarm, maintaining optimum internal movement as in~Figure \ref{methods:StabilityNullVector} is highly unlikely. The agent pairs may fluctuate between the 4 states~(Fig.~\ref{methods:Stability1}, \ref{methods:Stability2}, \ref{methods:Stability3} and \ref{methods:Stability4}); These transitions are \textit{jitter}. The degree to which this variation occurs can be measured using either the change in distance between the agents, or the change in the magnitude $P(b)$~(Equation~\ref{eq:CohesionRepulsion}) between the agents. Jitter arises as motion to maintain the structure of a swarm. A swarm coordination algorithm that reduces jitter is generally desirable.

\section{Modelling movement}\label{Section::Movement1}
Each agent within a swarm calculates its \textit{movement-direction vector}
based on its \textit{interaction} and \textit{destination} vectors. The
\textit{movement vector}~($b_{pos}$) is calculated using the unit
\textit{movement-direction vector}
(Equation~\ref{eq:InterAgentMovementDirectional1}) multiplied by the time
elapsed~($t$) in the system and the speed characteristic of the agent~($s_b$).

This process is carried out for every agent in the swarm to generate the
swarm's next position.

\begin{center}
\begin{equation}
\label{eq:AgentMovement}
b_{pos}=s_{b}t\big(v(b)\big)\string^ 
\end{equation}
\end{center}

The increment in the location of agent $b$ over time interval $t$ is shown in
Equation~\ref{eq:AgentMovement} where $s_b$ is the speed of agent~$b$. Note
that \string^ is the equivalent of $\hat{v} = \frac{v}{|v|}$ the normalised
vector.

\section{Swarm movement analysis\label{metric:MagnitudeDynamics2}}
The repulsion and cohesion vectors are generated for an agent through the
interaction of their field effects~(Fig.~\ref{methods:FieldEffects}). There are
a limited number of interactions that can occur; These are illustrated in
Fig.~\ref{methods:Stability1}, \ref{methods:Stability2},
\ref{methods:Stability3} and \ref{methods:Stability4}.

Following repeated experiments in the initial deployment of a swarm it was
found that using the same parameters for a swarm's environmental parameters and
the cohesion and repulsion ranges that the results where always very similar.
The results shown in this paper are therefore an example set typical of the
results obtained. The data extracts shown in Tables \ref{tab:SampleReplusion0},
\ref{tab:SampleReplusionPositive}, \ref{tab:SampleCohesionPositive} and
\ref{tab:SampleEquilibrium} are the simulation results that are produced by the
parameters listed in Table~\ref{tab:MetricPhysics1}. The simulation consists of
200 agents over a 20 second period. The cohesion of an agent pair is shown as
$k_cv_c$ and the repulsion as $k_rv_r$.

\begin{table}[H]
\begin{center}
\begin{tabular}{| p{2.5cm} | c | p{3cm} |}
\hline
\bf Weight \bf Component & \bf Setting & \bf Description \\ \hline
Sample Rate & 100 & ms - Unit sampling interval\\  \hline
$k_c$ & 5 & weight adjuster for cohesion bias\\  \hline
$k_r$ & 15 & weight adjuster for repulsion  bias\\  \hline
$k_d$ & 0 & weight adjuster for directional bias 0 for static baseline 100 from directional\\  \hline
Repulsion Boundary & 70 & units\\  \hline
Neighbour Distance & 80 & units\\  \hline
Speed & 20 & units/s\\  \hline
\end{tabular}\caption{Swarm model parameters} \label{tab:MetricPhysics1}
\end{center}
\end{table}

Figure~\ref{methods:Stability1} shows two agents within each other's cohesion
fields but sufficiently distant to be outside of the repulsion fields. In this
case $|k_cv_c| > 0$ and $|k_rv_r| = 0$: the result is the agent's resultant
magnitude $v(b)$ causes the agent to move towards its neighbour $b'$. Likewise
the neighbour's resultant magnitude $v(b')$ will cause it to move towards $b$.
Table~\ref{tab:SampleReplusion0} shows the repulsion magnitude with a value of
0. The only influence on the agent pair are cohesive vectors.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/Stability1}
{Internal movement cohesion (no repulsion).\label{methods:Stability1}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/Stability1}
%% \end{center}
%% \caption{Internal movement cohesion (no repulsion)} \label{methods:Stability1}
%% \end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l |}
\hline
Log &	Id &	N.Id &	Distance &	{\color{green}Cohesion} &	{\color{red}Repulsion} 	\\ \hline
0 &	1 &	3 	 & 70.50359 &	{\color{green}352.51799} &	{\color{red}0} \\ \hline
0 &	1 &	100 & 71.78005 &	{\color{green}358.90027} &	{\color{red}0} \\ \hline
0 &	1 &	151 & 78.33995 &	{\color{green}391.69979} &	{\color{red}0} \\ \hline
0 &	2 &	99  &	72.04066 &	{\color{green}360.20334} &	{\color{red}0} \\ 
\hline
\end{tabular}\caption{Data extract ($|k_rv_r| = 0$)} \label{tab:SampleReplusion0}
\end{center}
\end{table}

Figure~\ref{methods:Stability2} shows two agents close together with repulsion
dominating cohesion: $|k_cv_c|~<~|k_rv_r|$. The resultant vector directs the
agents away from each other.

Table~\ref{tab:SampleReplusionPositive} shows the repulsion magnitude with a
value greater than the cohesion magnitude.

\Figure[t!](topskip=0pt, botskip=0pt,
midskip=0pt)[width=6cm]{figures/Stability2} {Internal movement
repulsion.\label{methods:Stability2}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/Stability2}
%% \end{center}
%% \caption{Internal movement repulsion} \label{methods:Stability2}
%% \end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l |}
\hline
Log &	Id &	N.Id &	Distance &	{\color{green}Cohesion} &	{\color{red}Repulsion} 	\\ \hline
0 & 1 & 2 & 28.32522 & {\color{green}141.62612} & {\color{red}1544.86014} \\ \hline
0 & 1 & 6 & 41.48517 & {\color{green}207.42586} & {\color{red}721.71736} \\ \hline
0 & 1 & 7 & 35.26412 & {\color{green}176.32064} & {\color{red}1034.27101} \\ \hline
0 & 1 & 8 & 43.54503 & {\color{green}217.72518} & {\color{red}637.90759} \\
\hline
\end{tabular}\caption{Data extract ($|k_cv_c| < |k_rv_r|$)} \label{tab:SampleReplusionPositive}
\end{center}
\end{table}

Figure~\ref{methods:Stability3} shows two agents close together but with
cohesion vector magnitudes greater than the repulsion magnitudes
$|k_cv_c|~>~|k_rv_r|$. The resultant vector draws the agents together. The
magnitude of the resultant cohesion vector reduces due to the cancelling effect
of the repulsion vector. Table~\ref{tab:SampleCohesionPositive} shows a data
extract with the cohesion magnitude greater than repulsion.

\Figure[t!](topskip=0pt, botskip=0pt,
midskip=0pt)[width=6cm]{figures/Stability3} {Internal movement cohesion and
repulsion.\label{methods:Stability3}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/Stability3}
%% \end{center}
%% \caption{Internal movement cohesion and repulsion} \label{methods:Stability3}
%% \end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l |}
\hline
Log &	Id &	N.Id &	Distance &	{\color{green}Cohesion} & {\color{red}Repulsion} 	\\ \hline
0 & 1 & 5 &	64.17214 &	{\color{green}320.86072} &	{\color{red}95.35676} \\ \hline
0 & 1 & 9 &	63.88049 &	{\color{green}319.40248} &	{\color{red}100.58590} \\ \hline
0 & 1 & 95 & 65.61522 &	{\color{green}328.07613} &	{\color{red}70.16681} \\ \hline
0 & 1 & 152 & 63.10700 & {\color{green}315.5350} & {\color{red}114.68844} \\ 
\hline
\end{tabular}\caption{Data extract ($|k_cv_c| > |k_rv_r|$)} \label{tab:SampleCohesionPositive}
\end{center}
\end{table}

Figure~\ref{methods:Stability4} shows two agents close together with
$|k_cv_c|~=~|k_rv_r|$ the resultant vector is a \textit{null vector} and the
agents have no overall influence upon each other due to the magnitude of the
resultant vector being zero. Table~\ref{tab:SampleEquilibrium} is an extract of
data from the simulator. The data shows near equilibrium but due to the dynamic
nature of a swarm system no agents meet the condition fully.

\Figure[t!](topskip=0pt, botskip=0pt,
midskip=0pt)[width=6cm]{figures/Stability4} {Internal movement
equilibrium.\label{methods:Stability4}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=5cm]{figures/Stability4}
%% \end{center}
%% \caption{Internal movement equilibrium} \label{methods:Stability4}
%% \end{figure}

\begin{table}[H]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l |}
\hline
Log &	Id &	N.Id &	Distance &	{\color{green}Cohesion} &	{\color{red}Repulsion} 	\\ \hline
7 & 76 &	91 & 55.39031 & {\color{green}276.95156} & {\color{red}276.94684} \\ \hline
24 & 75 & 6 & 55.39032 & {\color{green}276.95160} & {\color{red}276.94661} \\ \hline
32 & 72 & 38 &	55.39002 & {\color{green}276.95013} & {\color{red}276.95370} \\ \hline
35 & 63 & 64 & 55.39022 &	{\color{green}276.95113} &	{\color{red}276.94887} \\
\hline
\end{tabular}\caption{Data extract ($|k_cv_c| \approx |k_rv_r|$)} \label{tab:SampleEquilibrium}
\end{center}
\end{table}

\section{Inter-Agent Distance metric}\label{Section:DistanceDynamics}
The inter-agent distance metric considers a swarm in terms of how the agents
are physically distributed: i.e. only the inter-agent distances and their
standard deviation are considered. The standard deviation indicates the extent
to which the swarm is out of balance and therefore the degree of movement
required to rebalance the swarm structure. If the standard deviation is zero
then all the agents are evenly spaced.

%The distance metric does not take into consideration the vector magnitudes between the agents as discussed above. The metric therefore is unable to identify the potential state of the swarm in terms of its cohesive or repulsive state.

Navarro and Fernando describe a \textit{mean distance error} metric that is
based on the variations in distances between inter-agent spaces~\cite{NIM:09}.
This is the same as the standard deviation of the distance-based internal
movement metric.

%% \section{Distance metric internal movement}
%% The distance based internal movement is measured by identifying the mean length of the vectors between an agent and its neighbours. As with the \textit{agent resultant magnitude} a coordination algorithm produces `jitter' which is the variations from the mean. In the case of the distance based metric the jitter is identified by the changes in the distances rather than the changes in vector magnitude (\textit{agent resultant magnitude}). The distance metric is the mean and the standard deviation `jitter' of the inter-agent distances.

% \section{Calculating distance internal movement}
Equation~\ref{eq:SwarmStabilityDistance1} calculates the mean distance of an agent to its neighbours $nbr(b)$. 
%% The relative position vector generated for an agent $b$ to its neighbour $b'$, $bb'$, is shown in~(\ref{eq:FlyToCentre1}). The magnitude of that vector gives the distance between two agents. For an individual agent the average magnitude $\mu_d(b)$ is calculated as \ref{eq:SwarmStabilityDistance1} where $b$ is the agent and $|nbr(b)|$ is the number of neighbours.

\begin{equation}
\label{eq:SwarmStabilityDistance1}
\mu_d(b) = \frac{1}{\mathlarger{|nbr(b)|}}\left({\mathlarger{\sum_{b' \in nbr(b)}}|bb'|}\right)
\end{equation}

The mean distance for a swarm is calculated
by~Equation~\ref{eq:SwarmStabilityDistance2}. All the inter-agent distances are
included for the swarm ($S$).

%% $\sum_{b \in S}|nbr(b)|$ calculates how many inter-agent relationships exist in the swarm and $\sum_{b' \in nbr(b)}|bb'|$ calculates the total distance between each agent and its neighbours. $\sum_{b \in S}$ iterates over all the agents in the swarm~($S$).

\begin{equation}
\label{eq:SwarmStabilityDistance2}
\mu_d(S) = \frac{\mathlarger{\sum_{b \in S}}~\mathlarger{\sum_{b' \in nbr(b)}}|bb'|}{\mathlarger{\sum_{b \in S}|nbr(b)|}}
\end{equation}

%\section{Standard deviation in distance metric}\label{Section:VarianceInDistance}
The mean distance provides a indication of the large scale structure of the
swarm. However it is not sufficient to give an indication of the internal
distribution of the agents. The standard deviation clarifies the distribution
within the swarm as shown in~Equation~\ref{eq:SwarmStabilityDistance3}.

%% $(|bb'| - \mu(S))^2$ is the square of the difference in a distance to the mean and $\sum_{b \in S}~\sum_{b' \in nbr(b)}$ calculates the number of inter-agent interactions.

\begin{equation}
\label{eq:SwarmStabilityDistance3}
\sigma_d(S) = \sqrt{\frac{\mathlarger{\sum_{b \in S}}~\mathlarger{\sum_{b' \in nbr(b)}}\Big(|bb'| - \mu_d(S)\Big)^2}{\mathlarger{\sum_{b \in S}|nbr(b)|}}}
\end{equation}

The distance-based metric for the internal distribution of the agents is
therefore $\mu_d(S)$ and $\sigma_d(S)$. This can be written informally as:

\begin{equation}
\label{eq:SwarmDistanceMetric}
\psi_d(S) = \mu_d(S)\pm \sigma_d(S)
\end{equation}

\Figure[t!](topskip=0pt, botskip=0pt,
midskip=0pt)[width=14cm]{figures/BaselineDistance1} {Baseline internal movement
- distance.\label{coord:BaselineDistance1}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=9cm]{figures/BaselineDistance1}
%% \end{center}
%% \caption{Baseline internal movement - distance\label{coord:BaselineDistance1}}
%% \end{figure}

\section{Cohesion/repulsion metric}\label{Section:MagnitudeDynamics}
% An agent's \textit{resultant magnitude} is identified by Gazi and Passino~\cite{GP:11} and Barnes et al~\cite{BFV:07} as a `resultant characteristic' of a swarm. The sum of the cohesion and repulsion vectors between an agent $b$ and a neighbour $b'$ produce a resultant vector.

% The cohesion/repulsion internal movement (\textit{cohesion/repulsion metric}) is measured by identifying the balance between the cohesion and repulsion vectors between agents as shown in~(\ref{eq:CohesionRepulsion}). This is derived from the \textit{inter-agent cohesion/repulsion vector} as shown in~(\ref{eq:InterAgentMovementDirectional1}). `Jitter' in the case of the agent resultant \textit{cohesion/repulsion metric} is measured as the deviation created by the inter-agent relationships. The identification of this deviation produces the clarifying component of the agent \textit{cohesion/repulsion metric}.
% %% The magnitude based metric identifies the resultant internal magnitude (vector magnitude) through the cohesive and repulsive magnitudes that exist between each agent pair in a swarm. 
% An agent's resultant vector magnitude is used as a scalar value to give an overall `size' to the agent relationship. The resultant scalar value is then modified based upon the `size' of the two relationship magnitudes.

% This paper will refer to the scalar resultant magnitude as the \textit{`cohesion/repulsion metric'} of the relationship. The `state' of a swarm is therefore the effect the environmental constraints and algorithms have upon the agent resultant magnitude. The agent's \textit{`cohesion/repulsion metric'} can be considered part of the `quality' measure for a swarm's performance.

% The resultant \textit{cohesion/repulsion metric}, which is a scalar, is generated by comparing the length of the cohesion and repulsion vector magnitudes. If the cohesion magnitude is larger than the repulsion vector magnitude the resultant length is considered to be positive. If the repulsion vector magnitude is larger than the cohesion vector magnitude the resultant length is considered to be negative.

% The \textit{`cohesion/repulsion metric'} is therefore the scalar value of the \textit{`agent resultant magnitude'} which is altered based upon the difference between the repulsion and cohesion magnitudes~(\ref{eq:CohesionRepulsion}).

This paper defines the cohesion/repulsion metric scalar for an agent as the
length of $v(b)$  as defined in Equation~\ref{eq:InterAgentMovement2}: that is,
the weighted effects of cohesion and repulsion by neighbours of the agent. This
number carries a negative sign if repulsion dominates and a positive sign if
cohesion dominates. This is summarised in Equation~\ref{eq:CohesionRepulsion}.

If an agent's {cohesion/repulsion metric} is negative the swarm's bias is to
expand as the repulsion is greater than cohesion. This is seen in the
disorganised stage of a swarm~(Fig.~\ref{coord:BaselineMagnitude1}). The
disorganised stage is the initial random deployment of the agents. If the agent
\textit{agent resultant magnitude} is positive then the swarm is exhibiting a
tendency to contract and this indicates the swarm is a cohesive entity. This
could also be described as the swarm being `sticky' as the agents bias is to
`pull' towards each other.

The \textit{cohesion/repulsion metric} on its own does not give a complete
measure of a swarm's internal state. There needs to be a qualifying component
to the metric that identifies the degree of deviation in the resultant
magnitude, this is the \textit{jitter}. The smaller the deviation the more
uniform the structure of the swarm. These two components identify the degree to
which a swarm has progressed towards a stable state.

The \textit{cohesion/repulsion metric} provides a view of the swarm's state
through the balance between the repulsive and the cohesive vectors that are
being applied to each agent. The deviation component identifies the degree to
which the swarm has stabilised. The ideal status for inter-agent interactions
would be for the agents to have a (\textit{cohesion/repulsion metric}) of zero
or above. This would indicate that the agents are distributed such that they
are at their distribution limit (outer most range of the cohesion field) or at
a level that causes the agents to `pull' together. The ideal degree of
deviation is zero as this indicates an even distribution of agents. These two
aspects of a swarm's features are not considered by Gazi and
Passino~\cite{GP:11} or Barnes et al~\cite{BFV:07} as a means of quantifying
the structure of a swarm in terms of stability.

% \section{Cohesion/repulsion internal movement model}\label{Section:StabilityModel}

Section~\ref{Section:StabilityMagnitude}~discussed calculation of resultant
cohesion~(Equation~\ref{eq:FlyToCentre1}) and resultant
repulsion~(Equation~\ref{eq:Repulsion1}) vectors for each agent deriving a
combined \textit{cohesion/repulsion vector} using as shown in
Equation~\ref{eq:InterAgentMovement2}.

%% This value represents the overall `potential' of an agent. This vector when normalised produces a component of the \textit{movement-destination vector} for a swarm. If the total cohesion/repulsion vector is zero (null vector) then the agent will not move. $M(b)$ is the total \textit{cohesion/repulsion vector} for agent $b$ defined by:
%% 
%% \begin{equation}
%% \label{eq:BotStabilityT}
%% M(b) = k_cv_c(b) + k_rv_r(b)
%% \end{equation}

Now, we define for each agent $b$ a \textit{cohesion/repulsion scalar}, $P(b)$,
to measure the influence of the neighbours of $b$ as cohesive (positive) or
repulsive (negative); shown in Equation~\ref{eq:CohesionRepulsion}.

\begin{equation}
\label{eq:CohesionRepulsion}
P(b) = \left\{\begin{array}{lll}
               |v(b)|& \mathrm{if} & |k_cv_c(b)| > |k_r v_r(b)|\\
              -|v(b)|& \mathrm{otherwise}
              \end{array}\right.
\end{equation}

Although it is possible for $v(b)$ to be null there could still be variation in
the constituent components. The variation calculation (standard deviation) as
shown in Equation~\ref{eq:SwarmStabilityQuotientT}.

%% Equation~\ref{eq:SwarmStabilityMetricT} is the mean of the cohesion/repulsion scalar for an agent.
%% 
%% \begin{equation}
%% \label{eq:SwarmStabilityMetricT}
%% \mu_p(b) = \frac{P(b)}{|nbr(b)|}
%% \end{equation}

%% \ref{eq:SwarmStabilityMetricT} should be considered as being applied at discrete points in time within a simulation. The formulae could also be shown based on time $t$ (\ref{eq:SwarmStabilityMetricT2}) where $b_t$ is an agent at a specific time interval. All formulae for agent resultant magnitude in this thesis will be considered as being applied at a point in time and will therefore be shown without a $t$ subscript.
%% 
%% \begin{equation}
%% \label{eq:SwarmStabilityMetricT2}
%% \mu_p(b_{t}) = \frac{P(b_{t})}{|nbr(b_{t})|}
%% \end{equation}

The mean cohesion/repulsion scalar for the swarm is now given
by~(\ref{eq:SwarmStabilityMetricT3}).

\begin{equation}
\label{eq:SwarmStabilityMetricT3}
\mu_p(S) = \frac{\mathlarger{\sum_{b \in S} P(b)}}{\mathlarger{\sum_{b \in S}}|nbr(b)| + \mathlarger{\sum_{b \in S}}|rep(b)|}
\end{equation}

%% \section{Standard deviation in agent cohesion/repulsion metric}\label{Section:VarianceInPotential}
%% To improve the metric clarification in terms of the deviation from the \textit{cohesion/repulsion metric} norm is calculated. The standard deviation of the entire swarm  is shown in Equation \ref{eq:SwarmStabilityMetricT}.

The standard deviation associated with this mean is calculated as
Equation~\ref{eq:SwarmStabilityQuotientT}.

%%  where $\sigma_p(S)$ is the standard deviation and $\mu_p(S)$ is the mean at the same point in time.

\begin{equation}
\label{eq:SwarmStabilityQuotientT}
\sigma_p(S) = \sqrt{\frac{\mathlarger{\sum_{b \in S}}\Big(P(b)-\mu_p(S)\Big)^2}{\mathlarger{\sum_{b \in S}}|nbr(b)| + \mathlarger{\sum_{b \in S}}|rep(b)|}}
\end{equation}

The metric for the internal movement is this pair of numbers, the mean and
standard deviation of the swarm's internal \textit{cohesion/repulsion}. The
pair $\mu_p(S)$, $\sigma_p(S)$ may be written informally as:

\begin{equation}
\label{eq:SwarmPotentialMagnitude}
\psi_p(S) = \mu_p(S)\pm \sigma_p(S)
\end{equation}

Figure~\ref{coord:BaselineMagnitude1} shows the evolution of the metric in time.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=14cm]{figures/BaselineMagnitude1}
{Baseline internal movement - cohesion/repulsion.\label{coord:BaselineMagnitude1}}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=9cm]{figures/BaselineMagnitude1}
%% \end{center}
%% \caption{Baseline internal movement - magnitude\label{coord:BaselineMagnitude1}}
%% \end{figure}

\section{Application: Area filling}\label{Section:ApplicationFloodFilling}

Area filling is a technique used to fill an enclosed area with agents such that
they are distributed as `effectively' as possible throughout an area.

Filling an area can be applied to tasks that require an unknown environment to
be analysed or surveyed. Consider a disaster area following a landslide or a
building collapse following an earthquake; the movement in the land and
buildings will produce areas that are unmapped. The unmapped areas may require
investigation to locate people, or resources, or to create some form of sensor
network to analyse the conditions within the area, such as creating a heat map
or identifying the location of toxic gases.

The concept of using a swarm to provide coverage over unknown areas is a
current area of research. Alvissalim et al. discuss the application of
commercially available drones to provide a communication infrastructure across
an unmapped disaster area~\cite{AZHMJJM:12}. Scheutz and Bauer (2006) use both
cohesion and repulsion as a mechanism to create coverage of an area that
requires protection in an adversarial environment~\cite{SB:06}.

The concept behind the area filling is to increase both the repulsion and
cohesion fields over a period of time. This increase in the field effects makes
the agents increase the distance between each other, expanding the swarm as a
whole. The expansion increases until, due to boundary compression, the swarm is
unable to expand further. In an extreme case the expansion will result in a set
of field effects that create a mesh-based swarm structure rather than the
desired hexagonal structure that the field effects should create. Identifying
the \textit{cohesion/repulsion metric} falling below zero indicates area
saturation has occurred.

To test this hypothesis a swarm is modelled in the simulator. The model
consists of an obstacle-based enclosed space and a swarm consisting of 60
agents. The experiments parameters for the simulation are shown
in~Table~\ref{tab:FillParameters1}

\begin{table}[H]
\begin{center}
\begin{tabular}{| p{1.8cm} | p{1.5cm} | p{4.0cm} |}
\hline
\bf Parameter & \bf Value  & \bf Description \\ \hline
$k_c$         & 5          & weight adjuster for cohesion bias\\ \hline
$k_r$         & 60         & weight adjuster for repulsion bias\\ \hline
Sample rate   & 100ms      & proximity sensor rate\\ \hline
Speed         & 20 units/s & agent speed\\ \hline
\end{tabular}\caption{Swarm parameters} \label{tab:FillParameters1}
\end{center}
\end{table}

The field effects are incremented in turn, cohesion range followed by repulsion
range. After each repulsion field change the swarm is allowed to redistribute
itself and stabilise. Table~\ref{tab:FillSequence} shows the field effect
settings that are used for the simulation. The field effects are selected so as
to ensure the swarm parameters have the potential to create a hexagonal swarm.
Figure \ref{emerge:FLOOD0} shows the initial swarm deployment.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD0}
{Flood fill Start.\label{emerge:FLOOD0}}

Figure~\ref{emerge:FLOOD1} shows the point a which data logging starts. The
swarm at this stage has expanded into the space but is not filling it. This
presents itself as a positive cohesion/repulsion metric value which indicates
the swarm is cohesive and the agents are still ``pulling" together.
Fig.~\ref{emerge:FLOOD6} shows the final swarm distribution for the simulation,
at this point the cohesion/repulsion metric has fell below zero indicating the
swarm is trying to expand.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD1}
{Data Capture Start.\label{emerge:FLOOD1}}

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=8cm]{figures/FLOOD6}
{Flood fill End.\label{emerge:FLOOD6}}

\begin{table}[H]
\begin{center}
\begin{tabular}{| p{1.8cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.3cm} | p{0.7cm} |}
\hline
\bf Weight \bf component & \bf 1 & \bf 2 & \bf 3 & \bf 4 & \bf 5 & \bf 6 & \bf 7 & \\ \hline
Repulsion Boundary & 50 & -  & 60 & -  & 70 & -  & 80 & units\\  \hline
Neighbour Distance & 60 & 70 & -  & 80 & -  & 90 & -  & units\\  \hline
\end{tabular}\caption{Field effect expansion sequence} \label{tab:FillSequence}
\end{center}
\end{table}

Figure~\ref{emerge:FIELDFILL-MAG} shows the cohesion/repulsion metric for the
swarm for the entire simulation. Between 100 seconds and 105 seconds there is a
significant change in the magnitude where the value becomes negative. This
indicates there is a compression effect on the swarm which indicates the area
is filled.

\Figure[t!](topskip=0pt, botskip=0pt,
midskip=0pt)[width=14cm]{figures/FIELDFILL-MAG} {Cohesion/Repulsion metric
0-120 seconds.\label{emerge:FIELDFILL-MAG}}

% \begin{figure}[H]
% \begin{center}
% \includegraphics[width=12cm]{figures/FIELDFILL-MAG}
% \end{center}
% \caption{Magnitude metric 0-120 seconds\label{emerge:FIELDFILL-MAG}}
% \end{figure}

Figure~\ref{emerge:FIELDFILL-DIST}~shows the distance metric for the same
simulation. The initial deployment is shown at 0 seconds followed by
incremental changes in the distances and variance of the distances as each
increment is made in the field effects. The simulation terminates after 120
seconds. There is no feature of this graph which indicates that the area is
filled. The cohesion/repulsion metric clearly tells us more.

\Figure[t!](topskip=0pt, botskip=0pt,
midskip=0pt)[width=14cm]{figures/FIELDFILL-DIST} {Distance metric 0-120
seconds.\label{emerge:FIELDFILL-DIST}}

% \begin{figure}[H]
% \begin{center}
% \includegraphics[width=12cm]{figures/FIELDFILL-DIST}
% \end{center}
% \caption{Distance metric 0-120 seconds\label{emerge:FIELDFILL-DIST}}
% \end{figure}

\section{Conclusion}\label{Section:MagnitudeDistanceComparison}
%% The two metrics (distance and \textit{cohesion/repulsion}) both measure the structure of a swarm in terms of the distribution of the agents and deviations from a mean. The \textit{cohesion/repulsion} metric also allows the state of the swarm to be identified.
%% 
%% The distance metric is based upon the physical \textit{distribution} of the agents and the cohesion/repulsion metric is based upon the logical \textit{interaction} of the agents.

Distance-based metrics measure a swarm's structure in terms of the position of
agents. This is independent of the algorithm that produces the swarm's
structure. It provides an analysis of agent distribution at a point in time but
does not allow the identification of the future behaviour of a swarm as
`expanding', `stable' or `contracting'.

\Figure[t!](topskip=0pt, botskip=0pt, midskip=0pt)[width=6cm]{figures/FieldEffects2}
{Heterogeneous field effects.\label{additional:fieldsWork}}

The new \textit{cohesion/repulsion} metric is based upon the algorithms that
have produced the structure and is therefore an analysis of how the algorithms
are performing rather than the structure presented. The new metric also
provides an indication of a swarm's potential future movement. This is
independent of the physical distribution of the agents within the swarm. The
metric also provides the same information as the distance metric in terms of
agent distribution when the field effects are consistent across all agents i.e.
monolithic field effects. This is shown in Fig. \ref{emerge:FIELDFILL-MAG} and
\ref{emerge:FIELDFILL-DIST} by the profiles of the two graphs being the same.
In addition the new metric provides a mechanism for evaluating swarms that use
different algorithms to create a swarming effect and allows the algorithms
inter-agent interactions to be compared rather than the resultant structure the
algorithm creates.

The two metrics can be used together to provide a deeper evaluation of a
swarm's structure and `potential'. Consider the following scenario: the
repulsion field is increased but the internal distances do not change; as a
result the \textit{cohesion/repulsion metric} falls below zero. This indicates
`something' is confining the swarm's distribution as discussed in Section
\ref{Section:ApplicationFloodFilling}. This analysis could be used in
identifying effective swarm distribution for the coverage of a sensor array as
discussed by Ramaithitima et al.~\cite{RWBK:15}

%% \begin{figure}[H]
%% \begin{center}
%% \includegraphics[width=6cm]{figures/FieldEffects2}
%% \end{center}
%% \caption{Complex swarm interactions\label{additional:fieldsWork}}
%% \end{figure}

\section{Future work}\label{Section:FutureWork}

The new metric can be applied to the analysis of the effects of coordination
algorithms upon a swarm's structure. An application to area filling has been
discussed~(Section~\ref{Section:ApplicationFloodFilling}). The metric can also
be applied to identifying the effects of `healing' algorithms, e.g. to remove
`voids' from a swarm. We plan to investigate this, and other, practical
applications of the metric.

The detection of a swarm's perimeter~\cite{MD:09, MJ:08, ZAPS:07, JG:13}
identifies a subset of a swarm which can be used for coordination.  Perimeter
detection is widely used in swarm coordination~\cite{ZFG:13, AKK:08,
APZDAMC:09, AZDPS:11}. A possible adaptation of this area of research is to
investigate the reduction of the field effects (cohesion and repulsion) on the
perimeter so as to increase the agent density in this region of the swam. This
change in the swarm's structure could improve the granularity of analysis, e.g.
when using sensor arrays at the frontier of the swarm.  However, this approach
leads to swarms of agents having heterogeneous field effects, as shown in
Fig.~\ref{additional:fieldsWork}. Clearly, the distance-based metric is not
useful in the analysis of such swarms. We believe that the cohesion/repulsion
metric introduced in this paper can identify a swarm's structure even when the
swarm is heterogeneous, since the metric is based directly upon inter-agent
field effects rather than the physical distribution of agents. This will also
be the subject of forthcoming work.

\clearpage

\bibliographystyle{plain}
\bibliography{thesis}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/eliot}}]{Neil Eliot} has been a Senior Lecturer at Northumbria University since 1998. Prior to this Neil worked in the chemical industry and the NHS. He was awarded his Ph.D. from Northumbria in 2017, his Pg.C. in 1994, and his B.Sc. in 1989 all in Computing. His research interests are in the area of swarm theory and cybersecurity. 
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/kenda}}]{David Kendall}
received the M.A. degree in literae humaniores from New College, Oxford University, and the M.Sc. and Ph.D. degrees in computing science from Newcastle University. He was a Research Associate at Newcastle University from 1987 to 1989, following experience as a Senior Software Engineer in industry from 1983 to 1986. He has been a Senior Lecturer in computing at Northumbria University since 1989, except for a brief period as a Lecturer in computer science at Durham University from 2001 to 2002. His research interests are in the areas of formal methods, embedded systems, distributed systems, security, and privacy.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Images/brock}}]{Michael Brockway}
has been a Senior Lecturer at Northumbria University for 17 years. His first degree is in pure and applied mathematics (1973-4), his Masters in mathematics (mathematical logic, category theory, 1975), and he has a PhD in computer science (formal methods for distributed embedded systems, 2010) from Northumbria University. Before that Michael worked in teaching and the computing industry.
\end{IEEEbiography}

\EOD

\end{document}